---
title: "Sarcoma data"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: united
    highlight: pygments
    df_print: paged
editor_options: 
  chunk_output_type: console
---

<style type="text/css">

.figure {
    margin-top: 100px;
    margin-bottom: 100px;
}

table {
    margin-top: 10px;
    margin-bottom: 25px !important;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      # fig.width = 7, fig.height = 5, 
                      fig.align='center'#, fig.show='hold'
                      )
options(gtsummary.print_engine = "gt")
options(gtsummary.as_gt.addl_cmds = "gt::tab_options(table.font.size = 14, data_row.padding = gt::px(1))")
```

```{r library}
library(tidyverse)
library(data.table)
library(VennDiagram)
library(lubridate)
library(gtsummary)
library(viridis)
```

```{r}
sarcoma_dna <- read_rds("/Users/colinccm/Documents/GitHub/Gillis/sarcoma_ch/sarcoma_dna.rds")
sarcoma_info <- read_rds("/Users/colinccm/Documents/GitHub/Gillis/sarcoma_ch/sarcoma_info.rds")
Chemot <- read_rds("/Users/colinccm/Documents/GitHub/Gillis/sarcoma_ch/Chemot.rds")
```

# How many sarcoma patients have samples from more than one time/date?
Are you interested to know if the sample is before or after chemotherapy drug?

```{r}
sarcoma_dna %>% 
  distinct(deidentified_patient_id, specimen_collection_date, .keep_all = TRUE) %>% 
  group_by(deidentified_patient_id) %>% 
  mutate(sample_count = factor(row_number(deidentified_patient_id))) %>% 
  ungroup() %>% 
  arrange(desc(sample_count)) %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  select(sample_count) %>% 
  tbl_summary()
```
`r sarcoma_dna %>% distinct(deidentified_patient_id, specimen_collection_date, .keep_all = TRUE) %>% group_by(deidentified_patient_id) %>% mutate(sample_count = row_number(deidentified_patient_id)) %>% ungroup() %>% filter(sample_count > 1) %>%   distinct(deidentified_patient_id, .keep_all = TRUE) %>% nrow()` patients have multiple samples.

# Summary for Histology of patients who have samples from more than one time.
Multiple histology per patients will be listed if a patient had multiple cancers. For example, "85003 invasive carcinoma of no special type (c50._)" and "88903 leiomyosarcoma nos" for one of a patient I saw. This is why we have a total number of histology in the table = 779 at and not 569.
```{r}
multiple_samples_patient <- sarcoma_dna %>% 
  distinct(deidentified_patient_id, specimen_collection_date, .keep_all = TRUE) %>% 
  group_by(deidentified_patient_id) %>% 
  mutate(sample_count = row_number(deidentified_patient_id)) %>% 
  ungroup() %>% 
  filter(sample_count > 1)
  
multiple_samples_patient <- paste(multiple_samples_patient$deidentified_patient_id, collapse = "|")

multiple_samples_dataset <- sarcoma_info %>% 
  filter(str_detect(deidentified_patient_id, multiple_samples_patient)) %>% 
  left_join(., sarcoma_dna, 
            by = "deidentified_patient_id")

multiple_samples_dataset %>% 
  select(histology) %>% 
  tbl_summary(sort = list(everything() ~ "frequency"))
```

# Summary for Chemotherapy Drug of patients who have samples from more than one time
The drug summary include all the regimen drugs that patients received between their first and last samples.  

```{r}
sarcoma_dna %>% 
  filter(str_detect(deidentified_patient_id, multiple_samples_patient)) %>% 
  left_join(., Chemot,
            by = "deidentified_patient_id") %>% 
  arrange(deidentified_patient_id, specimen_collection_date) %>% 
  group_by(deidentified_patient_id) %>% 
  mutate(first_specimen_date = first(specimen_collection_date)) %>% 
  mutate(last_specimen_date = last(specimen_collection_date)) %>% 
  ungroup() %>% 
  distinct(deidentified_patient_id, treatment_start_date, treatment_end_date, treatment, first_specimen_date, last_specimen_date) %>% 
  mutate(drugs_between_first_last_sample = case_when(
    treatment_start_date >= first_specimen_date &
      treatment_start_date < last_specimen_date        ~ "Yes"
  )) %>% 
  filter(drugs_between_first_last_sample == "Yes") %>% 
  select(treatment) %>% 
  tbl_summary(sort = list(everything() ~ "frequency"))
```

 It doesnt' look like a lot of drugs because our data is de-identified and we have `r sarcoma_dna %>% filter(str_detect(deidentified_patient_id, multiple_samples_patient)) %>% left_join(., Chemot, by = "deidentified_patient_id") %>% filter(treatment == "unk drug, 1800 date") %>% distinct(deidentified_patient_id, .keep_all = TRUE) %>% select(treatment) %>% nrow()` patients with at least one unknown date of drug and are excluded from the table above.  
```{r}
# sarcoma_dna %>% 
#   filter(str_detect(deidentified_patient_id, multiple_samples_patient)) %>% 
#   left_join(., Chemot,
#             by = "deidentified_patient_id") %>% 
#   select(treatment) %>% 
#   tbl_summary(sort = list(everything() ~ "frequency"))
# 
# sarcoma_dna %>%
#   filter(str_detect(deidentified_patient_id, multiple_samples_patient)) %>%
#   left_join(., Chemot,
#             by = "deidentified_patient_id") %>%
#   filter(treatment == "unk drug, 1800 date") %>% distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
#   select(treatment) %>% nrow()
```

```{r}
unk_data <- sarcoma_dna %>%
  filter(str_detect(deidentified_patient_id, multiple_samples_patient)) %>%
  left_join(., Chemot,
            by = "deidentified_patient_id") %>%
  filter(treatment == "unk drug, 1800 date")

unk_data_patient <- paste(unk_data$deidentified_patient_id, collapse = "|")


known_data <- sarcoma_dna %>%
  filter(str_detect(deidentified_patient_id, multiple_samples_patient)) %>% 
  left_join(., Chemot,
            by = "deidentified_patient_id") %>%
  filter(!str_detect(deidentified_patient_id, unk_data_patient)) %>% 
  arrange(deidentified_patient_id, specimen_collection_date) %>% 
  group_by(deidentified_patient_id) %>% 
  mutate(first_specimen_date = first(specimen_collection_date)) %>% 
  mutate(last_specimen_date = last(specimen_collection_date)) %>% 
  ungroup() %>% 
  distinct(deidentified_patient_id, treatment_start_date, treatment_end_date, treatment, first_specimen_date, last_specimen_date) %>% 
  mutate(drugs_between_first_last_sample = case_when(
    treatment_start_date >= first_specimen_date &
      treatment_start_date < last_specimen_date        ~ "Yes"
  )) %>% 
  mutate(interval_between_samples = interval(start = first_specimen_date, end = last_specimen_date)/
           duration(n=1, units = "days")) %>% 
  filter(drugs_between_first_last_sample == "Yes")

known_data %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  select(interval_between_samples) %>% 
  tbl_summary(statistic = list(all_continuous() ~ "{mean} ({min}, {max})"))
```

If we only include patients with multiple samples for whose we also know ALL their chemotherapy drug information, we can only include `r known_data %>% distinct(deidentified_patient_id) %>% nrow()` patients.  
Here would be their info on drugs between their first-last samples and histology.

```{r}

known_data %>% 
  select(treatment) %>% 
  tbl_summary(sort = list(everything() ~ "frequency"))
  
known_data_patient <- paste(known_data$deidentified_patient_id, collapse = "|")

multiple_samples_dataset %>% 
  filter(str_detect(deidentified_patient_id, known_data_patient)) %>%
  select(histology) %>% 
  tbl_summary(sort = list(everything() ~ "frequency"))
```

Let me know if you want anything alse or if you want to include the patients who have unknown drug date.




