---
title: "Sarcoma cancer and CH in Avatar data"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: united
    highlight: pygments
    df_print: kable
editor_options: 
  chunk_output_type: console
---

<style type="text/css">

.figure {
    margin-top: 100px;
    margin-bottom: 100px;
}

table {
    margin-top: 10px;
    margin-bottom: 25px !important;
}

th, td { padding: 5px; }

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      # fig.width = 7, fig.height = 5, 
                      fig.align='center'#, fig.show='hold'
                      )
options(gtsummary.print_engine = "gt")
options(gtsummary.as_gt.addl_cmds = "gt::tab_options(table.font.size = 14, data_row.padding = gt::px(1))")
```

```{r library}
library(tidyverse)
library(data.table)
library(VennDiagram)
library(lubridate)
library(gtsummary)
library(viridis)
library(ggforce)
library(kableExtra)
```

```{r}
sarcoma_patients <- 
  read_rds("/Users/colinccm/Documents/GitHub/Gillis/sarcoma_ch/sarcoma_patients limited blood samples type.rds")

treatment <- 
  read_rds("/Users/colinccm/Documents/GitHub/Gillis/sarcoma_ch/treatment_long.rds") %>% 
  filter(treatment_type == "chemo")

Global_data <- 
  inner_join(sarcoma_patients %>% 
               distinct(mrn, specimen_collection_date, 
                        date_of_diagnosis1, sample_type,
                        .keep_all = TRUE), 
             treatment, by = "mrn") %>% 
  # Create deidentified IDs
  mutate(rad = "sarcoma_study_") %>%
  group_by(mrn) %>% 
  mutate(id = cur_group_id()) %>%
  ungroup() %>%
  mutate(zero = 6 - nchar(id)) %>%
  mutate(ii = stringi::stri_dup("0", zero)) %>%
  select(c(rad, ii, id, mrn, everything())) %>%
  unite(deidentified_patient_id, rad:id, sep = "") %>% 
  select(c(deidentified_patient_id, mrn, everything(), -zero))

blood_patients <- Global_data %>% 
  # filter to patients who have blood samples
  filter(!is.na(specimen_collection_date))

write_rds(blood_patients, "blood_patients_long.rds")
```


```{r}
topo_drugs <- "epirubicin|etoposide|doxorubicin|daunorubicin|teniposide|mitoxantrone|amsacrine|actinomycin|adriamycin"
alk_drugs <- "melphalan|cyclophosphamide|nitrogen mustard|chlorambucil|busulfan|carboplatin|cisplatin|dacarbazine|procarbazine|carmustine|mitomycin|thiotepa|lomustine"
```


```{r}
blood_patients1 <- blood_patients %>%
  # mutate(first_chemo_date = first(treatment_start_date)) %>%
  # mutate(last_chemo_date = last(treatment_start_date)) %>%
  distinct() %>% 
  # CHEMO----
  # Select sample before 1st chemo
  mutate(blood_bf_chemo = case_when(
    specimen_collection_date <= treatment_start_date &
      treatment_line == "chemo_1"                                        ~ "Yes"
  )) %>%
  mutate(have_good_presample_chemo = blood_bf_chemo,
         presample_date_chemo = case_when(
           blood_bf_chemo == "Yes"                                       ~ specimen_collection_date
         ),
         presample_id_chemo = case_when(
           blood_bf_chemo == "Yes"                                       ~ sample_id
         )) %>% 
  mutate(days_from_presample_to_chemo = case_when(
    blood_bf_chemo == "Yes"                                             ~ interval(
      start = specimen_collection_date, end = treatment_start_date) /
      duration(n= 1, units = "days")
  )) %>% 
  group_by(mrn) %>% 
  fill(have_good_presample_chemo, presample_date_chemo,
       days_from_presample_to_chemo, presample_id_chemo,
       .direction = "updown") %>%
  ungroup() %>% 
  # Select sample after 1st chemo
  mutate(blood_af_chemo = case_when(
    specimen_collection_date > treatment_end_date &
      treatment_line == "chemo_1"                                        ~ "Yes"
  )) %>%
  mutate(postsample_id_chemo = case_when(
    blood_af_chemo == "Yes"                                               ~ sample_id
  )) %>% 
  mutate(postsample_date_chemo = case_when(
    blood_af_chemo == "Yes"                                               ~ specimen_collection_date
  )) %>% 
  
  # Topoisomerase II----
  # pre sample
  mutate(blood_bf_topoII = case_when(
    specimen_collection_date <= treatment_start_date &
      str_detect(treatment, topo_drugs)                                  ~ "Yes"
  )) %>%
  group_by(mrn) %>% 
  mutate(have_good_presample_topo = blood_bf_topoII,
         presample_date_topo = case_when(
           blood_bf_topoII == "Yes"                                      ~ last(specimen_collection_date)
         ),
         presample_id_topo = case_when(
           blood_bf_topoII == "Yes"                                      ~ last(sample_id)
         )) %>% 
  mutate(days_from_presample_to_topo = case_when(
    !is.na(presample_id_topo)                                            ~ interval(
      start = presample_date_topo, end = treatment_start_date) /
      duration(n= 1, units = "days")
  )) %>% 
  fill(have_good_presample_topo, presample_date_topo,
       days_from_presample_to_topo, presample_id_topo,
       .direction = "updown") %>%
  
  # post sample
  mutate(blood_af_topoII = case_when(
    specimen_collection_date > treatment_start_date &
      specimen_collection_date > treatment_end_date &
      str_detect(treatment, topo_drugs)                                  ~ "Yes"
  )) %>%
  mutate(blood_during_topoII = case_when(
    specimen_collection_date > treatment_start_date &
      specimen_collection_date < treatment_end_date &
      str_detect(treatment, topo_drugs)                                  ~ "Yes"
  )) %>%
  mutate(postsample_id_topo = case_when(
    blood_af_topoII == "Yes" |
      blood_during_topoII == "Yes"                                       ~ sample_id
  )) %>% 
  mutate(postsample_date_topo = case_when(
    blood_af_topoII == "Yes" |
      blood_during_topoII == "Yes"                                       ~ specimen_collection_date
  )) %>% 
  ungroup() %>% 
  mutate(days_from_drugsend_to_after_sample_topo = case_when(
    blood_af_topoII == "Yes" |
      blood_during_topoII == "Yes"                                      ~ interval(
        start = treatment_end_date, end = postsample_date_topo) /
      duration(n= 1, units = "days")
  )) %>% 
  select(mrn, deidentified_patient_id, specimen_collection_date, 
         blood_bf_topoII, blood_af_topoII, blood_during_topoII, treatment, everything()) %>% 
  
  # Alkylant agents
  mutate(blood_bf_alk = case_when(
    specimen_collection_date <= treatment_start_date &
      str_detect(treatment, alk_drugs)                                  ~ "Yes"
  )) %>%
  group_by(mrn) %>% 
  mutate(have_good_presample_alk = blood_bf_alk,
         presample_date_alk = case_when(
           blood_bf_alk == "Yes"                                        ~ last(specimen_collection_date)
         ),
         presample_id_alk = case_when(
           blood_bf_alk == "Yes"                                        ~ last(sample_id)
         )) %>% 
  mutate(days_from_presample_to_alk = case_when(
    !is.na(presample_id_alk)                                            ~ interval(
      start = presample_date_alk, end = treatment_start_date) /
      duration(n= 1, units = "days")
  )) %>% 
  fill(have_good_presample_alk, presample_date_alk,
       days_from_presample_to_alk, presample_id_alk,
       .direction = "updown") %>%
  mutate(blood_af_alk = case_when(
    specimen_collection_date > treatment_start_date &
      specimen_collection_date > treatment_end_date &
      str_detect(treatment, alk_drugs)                                  ~ "Yes"
  )) %>%
  mutate(postsample_id_alk = case_when(
    blood_af_alk == "Yes"                                               ~ sample_id
  )) %>% 
  mutate(postsample_date_alk = case_when(
    blood_af_alk == "Yes"                                               ~ specimen_collection_date
  )) %>% 
  ungroup() %>%
  
  mutate(days_from_drugsend_to_after_sample_alk = case_when(
      blood_af_alk == "Yes"                                             ~ interval(
      start = treatment_end_date, end = postsample_date_alk) /
      duration(n= 1, units = "days")
  )) %>% 
  
  mutate(duration_of_treatment = interval(
      start = treatment_start_date, end = treatment_end_date) /
      duration(n= 1, units = "days"),
      duration_of_treatment = case_when(
        duration_of_treatment == 0             ~ 1,
        TRUE                                   ~ duration_of_treatment
      )) %>% 
  
  # fill(blood_bf_topoII, 
  #      # blood_af_chemo,
  #      # blood_af_topoII,
  #      blood_bf_alk, 
  #      # blood_af_alk,
  #      # days_from_presample_to_drugs, days_from_drugs_to_after_sample,
  #      .direction = "updown") %>%
select(mrn, deidentified_patient_id, specimen_collection_date, 
         blood_bf_alk, blood_af_alk, treatment, everything()) %>% 
  arrange(mrn, deidentified_patient_id, specimen_collection_date, treatment_start_date)
```

<!-- # Treatment received by Sarcoma patients -->
<!-- ```{r} -->
<!-- treatment %>%  -->
<!--   filter(treatment_type != "radioT") %>%  -->
<!--   ungroup() %>%  -->
<!--   select(treatment) %>%  -->
<!--   tbl_summary() -->
<!-- ``` -->
<!-- <br> -->
<!-- <br> -->




# Patients with samples after or during treated with Topoisomerase II inhibitors (22 patients)

Topoisomerase II inhibitors include etoposide, doxorubicin, daunorubicin, teniposide, epirubicin, mitoxantrone, amsacrine, actinomycin, adriamycin.
```{r}
samples_topo <- blood_patients1 %>% 
  filter(have_good_presample_topo == "Yes" | blood_af_topoII == "Yes" | blood_during_topoII == "Yes") %>%
  # Excluded patients by checking medical records (treatment not for sarcoma)
  filter(!deidentified_patient_id %in% 
           c("sarcoma_study_000147", "sarcoma_study_000156", "sarcoma_study_000341",
             "sarcoma_study_000420", "sarcoma_study_000536",
             "sarcoma_study_000559", "sarcoma_study_000861")) %>% 
  mutate(group = "topo") %>% 
  select(mrn, deidentified_patient_id, group,
         date_of_diagnosis1, sample_type, 
         have_good_presample_topo, presample_id_topo, presample_date_topo, 
         days_from_presample_to_topo, 
         blood_during_topoII, blood_af_topoII, 
         postsample_id_topo, postsample_date_topo,
         days_from_drugsend_to_after_sample_topo, treatment, 
         treatment_start_date, treatment_end_date, duration_of_treatment) %>% 
  # group_by(mrn, deidentified_patient_id,
  #        date_of_diagnosis1, sample_type, have_good_presample_topo, 
  #        presample_id_topo, presample_date_topo, blood_during_topoII, 
  #        blood_af_topoII, postsample_id_topo, postsample_date_topo,
  #        days_from_presample_to_topo, days_from_drugsend_to_after_sample,
  #        treatment_start_date, duration_of_treatment
  #        ) %>%
  # summarise_at(vars(treatment, treatment_end_date), str_c, collapse = "!") %>%
  # ungroup() %>%
  # separate(treatment, paste("regimen", 1:2, sep = "_"),
  #          sep = "!", remove = TRUE,
  #          extra = "warn", fill = "right") %>%
  rename(have_good_presample_drug = have_good_presample_topo,
         presample_id_drug = presample_id_topo, presample_date_drug = presample_date_topo,
         blood_during_drug = blood_during_topoII, 
         blood_af_drug = blood_af_topoII, 
         postsample_id_drug = postsample_id_topo, postsample_date_drug = postsample_date_topo,
         days_from_drugsend_to_after_sample = days_from_drugsend_to_after_sample_topo
         ) %>% 
  # Keep the closest blood after
  distinct(mrn, deidentified_patient_id, group, blood_during_drug, blood_af_drug, .keep_all = TRUE) %>% 
  arrange(mrn, deidentified_patient_id, treatment_start_date)

# write_csv(samples_topo, "list of sarcoma samples topo.csv")
write_csv(samples_topo, "list of sarcoma patients with samples topo 06212022.csv")
```

<!-- # Patients with samples during treated with Topoisomerase II inhibitors (13 patients) -->
<!-- ```{r} -->
<!-- blood_patients %>%  -->
<!--   filter(blood_during_topoII == "Yes") %>%  -->
<!--   select(deidentified_patient_id, mrn,  treatment_line, treatment) %>%  -->
<!--   group_by(deidentified_patient_id) %>%  -->
<!--   summarise_at(vars(treatment), str_c, collapse = "!") %>%  -->
<!--   ungroup() %>%  -->
<!--   separate(treatment, paste("regimen", 1:10, sep = "_"),  -->
<!--            sep = "!", remove = TRUE,  -->
<!--            extra = "warn", fill = "right") %>% -->
<!--   purrr::keep(~!all(is.na(.))) -->
<!-- ``` -->
<!-- <br> -->
<!-- <br> -->

# Patients treated with Alkylating agents (64 patients)

Alkylating agents include melphalan, cyclophosphamide, nitrogen mustard, chlorambucil, busulfan, carboplatin, cisplatin, dacarbazine, procarbazine, carmustine, mitomycin, thiotepa, lomustine.
```{r}
samples_alk <- blood_patients1 %>%
  filter(have_good_presample_alk == "Yes" & blood_af_alk == "Yes") %>%
  # Excluded patients by checking medical records (treatment not for sarcoma)
  filter(!deidentified_patient_id %in% 
           c("sarcoma_study_000091", "sarcoma_study_000203", "sarcoma_study_000488", 
             "sarcoma_study_000539", "sarcoma_study_000864")) %>%  
  mutate(group = "alkylating") %>% 

  select(mrn, deidentified_patient_id, group,
         date_of_diagnosis1, sample_type, 
         have_good_presample_alk, presample_id_alk, presample_date_alk,
         days_from_presample_to_alk, 
         blood_af_alk, postsample_id_alk, postsample_date_alk,
         days_from_drugsend_to_after_sample_alk, treatment, 
         treatment_start_date, treatment_end_date, duration_of_treatment) %>% 
  # group_by(mrn, deidentified_patient_id, 
  #          date_of_diagnosis1, sample_type, have_good_presample_alk, 
  #          presample_id_alk, presample_date_alk,
  #          blood_af_alk, postsample_id_alk, postsample_date_alk,
  #          days_from_presample_to_alk, days_from_drugsend_to_after_sample, 
  #          treatment_start_date, duration_of_treatment) %>% 
  # summarise_at(vars(treatment, treatment_end_date), str_c, collapse = "!") %>% 
  # ungroup() %>% 
  # separate(treatment, paste("regimen", 1:2, sep = "_"), 
  #          sep = "!", remove = TRUE, 
  #          extra = "warn", fill = "right") %>% 
  # Keep the closest blood after
  rename(have_good_presample_drug = have_good_presample_alk,
         presample_id_drug = presample_id_alk, presample_date_drug = presample_date_alk,
         days_from_presample_to_drug = days_from_presample_to_alk, 
         blood_af_drug = blood_af_alk, 
         postsample_id_drug = postsample_id_alk, postsample_date_drug = postsample_date_alk,
         days_from_drugsend_to_after_sample = days_from_drugsend_to_after_sample_alk
         ) %>% 
  distinct(mrn, deidentified_patient_id, group, blood_af_drug, .keep_all = TRUE) %>% 
  arrange(mrn, deidentified_patient_id, treatment_start_date)

# write_csv(samples_alk, "list of sarcoma samples alkylating.csv")
write_csv(samples_alk, "list of sarcoma patients with samples alkylating 06212022.csv")
```


# Patients treated with ANY drugs

```{r}
# samples_chemo <- blood_patients %>%
#   filter(have_good_presample_chemo == "Yes" & blood_af_chemo == "Yes") %>%
#   # Excluded patients by checking medical records (treatment not for sarcoma)
#   filter(deidentified_patient_id %in%
#            c("sarcoma_study_000007", "sarcoma_study_000014", "sarcoma_study_000084",
#              "sarcoma_study_000130", "sarcoma_study_000229", "sarcoma_study_000792",
#              "sarcoma_study_000813")) %>%
#   # Exclude topo or alk
#   filter(!str_detect(deidentified_patient_id, paste0(samples_topo$deidentified_patient_id, collapse = "|"))) %>% 
#   filter(!str_detect(deidentified_patient_id, paste0(samples_alk1$deidentified_patient_id, collapse = "|"))) %>% 
#   
#   mutate(group = "other chemo") %>% 
#   select(mrn, 
#          deidentified_patient_id, 
#          date_of_diagnosis1,
#          sample_type, 
#          have_good_presample_chemo, 
#          presample_id_chemo, presample_date_chemo,
#          blood_af_chemo, 
#          postsample_id_chemo, postsample_date_chemo,
#          days_from_presample_to_chemo,
#          days_from_drugsend_to_after_sample, treatment, 
#          treatment_start_date, treatment_end_date,
#          duration_of_treatment) # %>% 
  # distinct(mrn, .keep_all = TRUE)

# write_csv(samples_chemo, "list of sarcoma samples chemo.csv")
# write_csv(samples_chemo, "list of sarcoma patients with samples chemo.csv")
```

```{r}
samples <- bind_rows(samples_topo, samples_alk) %>% 
  arrange(mrn, deidentified_patient_id, treatment_start_date)
write_csv(samples, "list of sarcoma patients with samples of interest 06212022.csv")
```



