---
title: "Sarcoma cancer and CH in Avatar data"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: united
    highlight: pygments
    df_print: kable
editor_options: 
  chunk_output_type: console
---

<style type="text/css">

.figure {
    margin-top: 100px;
    margin-bottom: 100px;
}

table {
    margin-top: 10px;
    margin-bottom: 25px !important;
}

th, td { padding: 5px; }

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      # fig.width = 7, fig.height = 5, 
                      fig.align='center'#, fig.show='hold'
                      )
options(gtsummary.print_engine = "gt")
options(gtsummary.as_gt.addl_cmds = "gt::tab_options(table.font.size = 14, data_row.padding = gt::px(1))")
```

```{r library}
library(tidyverse)
library(data.table)
library(VennDiagram)
library(lubridate)
library(gtsummary)
library(viridis)
library(ggforce)
library(kableExtra)
```

```{r}
sarcoma_patients <- 
  read_rds("/Users/colinccm/Documents/GitHub/Gillis/sarcoma_ch/sarcoma_patients.rds")

treatment <- 
  read_rds("/Users/colinccm/Documents/GitHub/Gillis/sarcoma_ch/treatment_long.rds") %>% 
  filter(treatment_type == "chemo")

Global_data <- 
  inner_join(sarcoma_patients, treatment, by = "mrn") %>% 
  # Create deidentified IDs
  mutate(rad = "sarcoma_study_") %>%
  group_by(mrn) %>% 
  mutate(id = cur_group_id()) %>%
  ungroup() %>%
  mutate(zero = 6 - nchar(id)) %>%
  mutate(ii = stringi::stri_dup("0", zero)) %>%
  select(c(rad, ii, id, mrn, everything())) %>%
  unite(deidentified_patient_id, rad:id, sep = "") %>% 
  select(c(deidentified_patient_id, mrn, everything(), -zero))

blood_patients <- Global_data %>% 
  # filter to patients who have blood samples
  filter(!is.na(specimen_collection_date))

write_rds(blood_patients, "blood_patients_long.rds")
```


```{r}
topo_drugs <- "epirubicin|etoposide|doxorubicin|daunorubicin|teniposide|mitoxantrone|amsacrine|actinomycin"
alk_drugs <- "melphalan|cyclophosphamide|nitrogen mustard|chlorambucil|busulfan|carboplatin|cisplatin|dacarbazine|procarbazine|carmustine|mitomycin|thiotepa|lomustine"
```


```{r}
blood_patients <- blood_patients %>%
  group_by(mrn) %>% 
  mutate(first_chemo_date = first(treatment_start_date)) %>% 
  mutate(last_chemo_date = last(treatment_start_date)) %>% 
  
  mutate(blood_bf_chemo = case_when(
    specimen_collection_date <= first_chemo_date                         ~ "Yes"
  )) %>%
  mutate(blood_af_chemo = case_when(
    specimen_collection_date > last_chemo_date                           ~ "Yes"
  )) %>%
  
  mutate(blood_bf_topoII = case_when(
    specimen_collection_date <= treatment_start_date &
      str_detect(treatment, topo_drugs)                                  ~ "Yes"
  )) %>%
  mutate(blood_af_topoII = case_when(
    specimen_collection_date > treatment_start_date &
      str_detect(treatment, topo_drugs)                                  ~ "Yes"
  )) %>%
  
  mutate(blood_bf_alk = case_when(
    specimen_collection_date <= treatment_start_date &
      str_detect(treatment, alk_drugs)                                  ~ "Yes"
  )) %>%
  mutate(blood_af_alk = case_when(
    specimen_collection_date > treatment_start_date &
      str_detect(treatment, alk_drugs)                                  ~ "Yes"
  )) %>%
  fill(blood_bf_chemo, blood_af_chemo, 
       blood_bf_topoII, blood_af_topoII,
       blood_bf_alk, blood_af_alk,
       .direction = "updown") %>% 
  ungroup() %>% 

  arrange(mrn, deidentified_patient_id, specimen_collection_date) %>%
  select(mrn, deidentified_patient_id, specimen_collection_date, 
         blood_bf_chemo, blood_af_chemo,
         blood_bf_topoII, blood_af_topoII,
         blood_bf_alk, blood_af_alk,
         treatment_start_date, treatment, 
         first_chemo_date, last_chemo_date,
         everything())
```

# Treatment received by Sarcoma patients
```{r}
treatment %>% 
  filter(treatment_type != "radioT") %>% 
  ungroup() %>% 
  select(treatment) %>% 
  tbl_summary()
```
<br>
<br>

# Patients treated with ANY drugs (53 patients)

```{r}
blood_patients %>% 
  filter(blood_bf_chemo == "Yes" & blood_af_chemo == "Yes") %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>%
  select(deidentified_patient_id) %>%
  purrr::keep(~!all(is.na(.)))
```


# Patients treated with Topoisomerase II inhibitors (2 patients)

Topoisomerase II inhibitors include etoposide, doxorubicin, daunorubicin, teniposide, epirubicin, mitoxantrone, amsacrine, actinomycin.
```{r}
blood_patients %>% 
  filter(blood_bf_topoII == "Yes" & blood_af_topoII == "Yes") %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>%
  select(deidentified_patient_id) %>%
  purrr::keep(~!all(is.na(.)))
```
<br>
<br>

# Patients treated with Alkylating agents (64 patients)

Alkylating agents include melphalan, cyclophosphamide, nitrogen mustard, chlorambucil, busulfan, carboplatin, cisplatin, dacarbazine, procarbazine, carmustine, mitomycin, thiotepa, lomustine.
```{r}
blood_patients %>% 
  filter(blood_bf_alk == "Yes" & blood_af_alk == "Yes") %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>%
  select(deidentified_patient_id) %>%
  purrr::keep(~!all(is.na(.)))
```




<!-- ```{r code pre and sequential} -->
<!-- blood_patients2 <- blood_patients %>%  -->
<!--   ###### Chemo samples ---- -->
<!--   mutate(had_good_sample_chemo = case_when( -->
<!--     blood_bf_chemo == "Yes"                    ~ "Yes" -->
<!--   )) %>% -->
<!--   group_by(deidentified_patient_id, had_good_sample_chemo) %>% -->
<!--   mutate(closest_chemo_presample_date = case_when( -->
<!--     had_good_sample_chemo == "Yes"            ~ last(specimen_collection_date) -->
<!--   )) %>% -->
<!--   mutate(interval_presample_chemo = case_when( -->
<!--     closest_chemo_presample_date == specimen_collection_date  ~  -->
<!--       (interval(start = closest_chemo_presample_date, end = chemotherapy_start_date_1) / -->
<!--                  duration(n = 1, units = "days")) -->
<!--   )) %>% -->
<!--   mutate(presample_id_chemo = case_when( -->
<!--     !is.na(interval_presample_chemo)       ~ sample_id -->
<!--   )) %>%  -->
<!--   mutate(presample_date_chemo = case_when( -->
<!--     !is.na(presample_id_chemo)                                     ~ specimen_collection_date -->
<!--   )) %>%  -->

<!--   # Find sequential samples -->
<!--   mutate(presample_chemo_within_limits = had_good_sample_chemo) %>%  -->
<!--   group_by(deidentified_patient_id) %>% -->
<!--   fill(presample_chemo_within_limits, #interval_presample_chemo, -->
<!--        closest_chemo_presample_date, .direction = "updown") %>% -->
<!--   mutate(seq_sample_chemo = case_when( -->
<!--     presample_chemo_within_limits == "Yes" & -->
<!--       blood_bf_chemo == "No" & -->
<!--       specimen_collection_date >= chemotherapy_end_date1_1 & -->
<!--       specimen_collection_date <= (chemotherapy_start_date_1 + days(365)) & -->
<!--       (specimen_collection_date <= radiation_start_date_1 | -->
<!--       is.na(radiation_start_date_1))                                ~ "Yes", -->
<!--     TRUE                                                            ~ "No" -->
<!--   )) %>% -->
<!--   group_by(mrn, seq_sample_chemo) %>%  -->
<!--   mutate(is_first_seq_sample = case_when( -->
<!--     seq_sample_chemo == "Yes" & -->
<!--     specimen_collection_date == first(specimen_collection_date)     ~ "Yes", -->
<!--     TRUE                                                            ~ "No" -->
<!--   )) %>%  -->
<!--   ungroup() %>%  -->

<!--   mutate(seqsample_id_chemo = case_when( -->
<!--     is_first_seq_sample == "Yes"                                    ~ sample_id -->
<!--   )) %>%  -->

<!--   mutate(seqsample_date_chemo = case_when( -->
<!--     !is.na(seqsample_id_chemo)                                     ~ specimen_collection_date -->
<!--   )) %>%  -->
<!--   mutate(seqsample_before_2nd_regimen = case_when( -->
<!--     !is.na(seqsample_id_chemo) & -->
<!--       specimen_collection_date <= chemotherapy_start_date_2        ~ "sample before 2nd regimen", -->
<!--     !is.na(seqsample_id_chemo) & -->
<!--       specimen_collection_date > chemotherapy_start_date_2         ~ "sample after" -->
<!--   )) %>%  -->
<!--   mutate(have_good_seqsample_chemo = case_when( -->
<!--     seq_sample_chemo == "Yes"                                       ~ "Yes" -->
<!--   )) %>%  -->
<!--   fill(have_good_seqsample_chemo, .direction = "updown") %>%  -->
<!--   mutate(time_chemo_seqsample = case_when( -->
<!--     seq_sample_chemo == "Yes"        ~ ( -->
<!--       interval(start = chemotherapy_start_date_1, end = specimen_collection_date) / -->
<!--         duration(n= 1, units = "days") -->
<!--     ))) %>% -->
<!--   mutate(interval_prepost_sample_chemo = case_when( -->
<!--     !is.na(seqsample_id_chemo)                                     ~ -->
<!--       abs(interval(start = closest_chemo_presample_date, end = specimen_collection_date) / -->
<!--                  duration(n = 1, units = "days")) -->
<!--   )) -->
<!-- ``` -->

<!-- # I. Data Exploration -->


<!-- # II. Table patients with Blood before and after treatment -->
<!-- ```{r} -->
<!-- tbl1_1 <- blood_patients2 %>%  -->
<!--   distinct(deidentified_patient_id, .keep_all = TRUE) %>%  -->
<!--   select(Chemotherapy = presample_chemo_within_limits) %>%  -->
<!--   tbl_summary(statistic = list(all_categorical() ~ "{n}")) %>%  -->
<!--   modify_header(update = list( -->
<!--   stat_0 ~ '**Patients with Blood before Treatment**' -->
<!-- )) -->


<!-- tbl1_2 <- blood_patients2 %>%  -->
<!--   filter(is.na(seqsample_id_Achemorad) & !is.na(seqsample_id_chemo)) %>% -->

<!--   distinct(deidentified_patient_id, .keep_all = TRUE) %>%  -->
<!--   select(Chemotherapy = seq_sample_chemo) %>%  -->
<!--   tbl_summary(statistic = list(all_categorical() ~ "{n}")) %>%  -->
<!--   modify_header(update = list( -->
<!--   stat_0 ~ '**Patients with Sequential Blood after Treatment(s)**' -->
<!-- )) -->

<!-- tbl_s1 <- tbl_merge(list(tbl1_1, tbl1_2)) -->

<!-- tbl_s1 %>% -->
<!--    modify_header(list(label ~ "**Treatment Type**")) %>%  -->
<!--   modify_spanning_header(everything() ~ NA_character_) %>% as_gt() %>%   -->
<!--   gt::tab_source_note(gt::md("*The number of patients with a blood sample before each treatment type is located  -->
<!--                              in the first column. <br>  -->
<!--                              From this number, the second column represent the number of patients with a blood  -->
<!--                              sample after each treatment type*")) -->

<!-- ``` -->

<!-- # III. Chemo group -->
<!-- ```{r} -->
<!-- chemo_patients <- blood_patients2 %>%  -->
<!--   filter(is.na(seqsample_id_Achemorad) & !is.na(seqsample_id_chemo)) -->

<!-- chemo_patients <- blood_patients2 %>% -->
<!--   filter( -->
<!--     str_detect( -->
<!--       deidentified_patient_id,  -->
<!--       paste0(chemo_patients$deidentified_patient_id, collapse = "|"))) -->
<!-- ``` -->



<!-- ```{r} -->
<!-- chemo_patients <- chemo_patients %>%  -->
<!--   group_by(deidentified_patient_id) %>%  -->
<!--   mutate(sequential_sample_count = factor(row_number(deidentified_patient_id))) %>%  -->
<!--   ungroup() %>%  -->
<!--   mutate(chemo_stop_T2 = interval(start = chemotherapy_end_date1_1, end = seqsample_date_chemo)/ -->
<!--              duration(n=1, units = "days")) %>%  -->
<!--   mutate(ID = dense_rank(deidentified_patient_id)) -->

<!-- df <- chemo_patients %>% -->
<!--   select(deidentified_patient_id, ID, -->
<!--          "number post treatment sample"  = sequential_sample_count,  -->
<!--          "days from pre sample to chemo start" = interval_presample_chemo, -->
<!--          chemo_stop_T2, -->
<!--          seqsample_before_2nd_regimen, -->
<!--          "days from chemo start to sequential sample" =  time_chemo_seqsample) -->

<!-- color.me1 <- which((df$ID %% 2)  == 1) -->
<!-- # color.me <- which(df$`number post treatment sample` == 2) -->
<!-- # color.me.bold <- which((df$ID %% 2)  == 1) -->
<!-- color.me.pre <- which(df$`number post treatment sample` == 1) -->
<!-- color.me.seq1 <- which(df$`number post treatment sample` == 2) -->


<!-- df %>% select(-ID) %>%  -->
<!--   kable(booktabs = T, align = "ccl") %>% -->
<!--   # kable_styling(latex_options = "striped", stripe_color = "red") #%>% -->
<!--   # row_spec(color.me, background = "#eee") %>% -->
<!--   row_spec(color.me1, background = "#eee") %>% -->
<!--   row_spec(color.me.pre, color = "blue") %>% -->
<!--   row_spec(color.me.seq1, color = "red") -->

<!-- df1 <- df %>% filter(!is.na(`days from pre sample to chemo start`)) %>%  -->
<!--   distinct(deidentified_patient_id, .keep_all = TRUE) %>%  -->
<!--   summarize(mean = round(mean(`days from pre sample to chemo start`),2), -->
<!--             median = round(median(`days from pre sample to chemo start`),2)) -->

<!-- df <- df %>% filter(!is.na(`days from chemo start to sequential sample`)) %>%  -->
<!--   distinct(deidentified_patient_id, .keep_all = TRUE) %>%  -->
<!--   summarize(mean = round(mean(`days from chemo start to sequential sample`),2), -->
<!--             median = round(median(`days from chemo start to sequential sample`),2)) -->
<!-- ``` -->


<!-- <br> -->

<!-- ## 5.List chemotherapy drugs (first regimen)  -->

<!-- ```{r} -->
<!-- chemo_patients %>%  -->
<!--   distinct(deidentified_patient_id, .keep_all = TRUE) %>%  -->
<!--   select(chemotherapy_drug_1) %>%  -->
<!--   tbl_summary(sort = list(everything() ~ "frequency")) -->

<!-- chemo_patients %>%  -->
<!--   distinct(deidentified_patient_id, .keep_all = TRUE) %>%  -->
<!--   filter(chemotherapy_drug_1 == "adriamycin; cyclophosphamide; paclitaxel") %>%  -->
<!--   select(chemotherapy_drug_1, line_gap_1) -->

<!-- chemo_patients %>%  -->
<!--   distinct(deidentified_patient_id, .keep_all = TRUE) %>%  -->
<!--   filter(chemotherapy_drug_1 == "5 fu; cyclophosphamide; epirubicin; docetaxel") %>%  -->
<!--   select(chemotherapy_drug_1, line_gap_1) -->
<!-- ``` -->

<!-- ## 6.Patients list -->
<!-- ```{r} -->
<!-- chemo_patients %>%  -->
<!--   distinct(deidentified_patient_id, .keep_all = TRUE) %>%  -->
<!--   select(deidentified_patient_id, age_at_diagnosis) -->

<!-- # chemo_patients %>% -->
<!-- #   select(deidentified_patient_id, -->
<!-- #          presample_id_chemo, -->
<!-- #          presample_date_chemo, -->
<!-- #          seqsample_id_chemo, -->
<!-- #          seqsample_date_chemo, -->
<!-- #          seqsample_before_2nd_regimen, -->
<!-- #          chemo_stop_T2, time_chemo_seqsample, -->
<!-- #          chemotherapy_start_date_1, chemotherapy_end_date1_1, -->
<!-- #          chemotherapy_drug_1) %>% -->
<!-- #   group_by(deidentified_patient_id) %>% -->
<!-- #   fill(everything(), .direction = "updown") %>% -->
<!-- #   distinct() %>% -->
<!-- #   kbl() %>% -->
<!-- #   kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% -->
<!-- #   scroll_box(width = "100%") -->
<!-- #  -->
<!-- # a <- chemo_patients %>% -->
<!-- #   filter(!is.na(presample_id_chemo) | !is.na(seqsample_id_chemo)) %>%  -->
<!-- #   select(mrn, deidentified_patient_id, -->
<!-- #          presample_id_chemo, -->
<!-- #          presample_date_chemo, -->
<!-- #          interval_presample_chemo, -->
<!-- #          seqsample_id_chemo, -->
<!-- #          seqsample_date_chemo, -->
<!-- #          time_chemo_seqsample, -->
<!-- #          chemotherapy_start_date_1, chemotherapy_end_date1_1, -->
<!-- #          chemotherapy_drug_1) %>% -->
<!-- #   group_by(deidentified_patient_id) %>% -->
<!-- #   fill(everything(), .direction = "updown") %>% -->
<!-- #   distinct() -->
<!-- # write_csv(a, "Samples list chemotherapy group in breast patients 03012022.csv") -->
<!-- ``` -->
